Notes for Week 12/30/19 - 1/3/19

### 2020-01-25 notes/weekly_notes/1.3.20.NotesForWeek.tex
updated weekly notes and project level notes

Suppose we observe independent samples $X = \set{x_1,\ldots,x_n}$ from a distribution $P$ with density $p$ supported on $\mathcal{X} = [0,1]^d$. Let $G_{n,r}$ be the random geometric graph of radius $r$ formed over the samples $X$, that is, the neighborhood graph with kernel $K(x) = \1(\norm{x} \leq r)$, and let $L$ denote the corresponding graph Laplacian matrix. In this week's notes, we will show that as long as $r$ is chosen to be sufficiently small, the eigenvalues of $L$ we care about scale at an appropriate rate. More precisely, we will prove the following Lemma.
\begin{lemma}
	\label{lem:higher_order_eigenvalue_tail_bound}
	Let $\kappa = n^{2d/(4s + d)}$. Suppose $p$ is uniformly bounded away from $0$ and $\infty$ on its support,
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Then there exists a constant $c$ such that for any $r \leq n^{-4/((2+d)(4s + d))}$, the eigenvalue $\lambda_{\kappa}$ is lower bounded,
	\begin{equation}
	\label{eqn:higher_order_eigenvalue_tail_bound}
	\lambda_{\kappa} \geq c\kappa^{2/d}nr^{2 + d}
	\end{equation}
	with probability $1 - o(n^{-1})$ as $n \to \infty$.
\end{lemma}
Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} establishes that eigenvalues of the graph Laplacian, suitably normalized, scale comparably to the eigenvalues of the continuum Laplacian operator $\mathcal{D}f := \textrm{div}(\nabla f)$. Along with an upper bound on the graph Sobolev semi-norm $R_{s,n}(f)$ (established in the \textit{12.20.19 Notes}), this allows us to argue that if $f$ is a smooth function with large $\mathcal{L}_2$ norm, then the projection of a function $f$ onto the eigenvectors $v_1,\ldots,v_{\kappa}$ is still quite large (measured in empirical norm); or, equivalently, the bias induced by projection onto these eigenvectors is small.

@@

The eigenvalues of lattices and complete graphs are known to satisfy, respectively
\begin{equation*}
\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
\end{equation*}
and by standard facts regarding the eigenvalues of tensor product graphs (\textcolor{red}{reference}), we have that the spectrum $\Lambda(H)$ satisfies
\begin{equation*}
\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
\end{equation*}
For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
\begin{align*}
\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
\end{align*}
where the last inequality can be verified by a quick calculation in light of $\kappa = n^{2d/(4s + d)}$ and $r \leq n^{-4/((2+d)(4s + d))}$. Since we've already shown $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ with probability $1 - o(n^{-1})$, this completes the proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}. 
\end{document}

>>>>

Suppose we observe independent samples $X = \set{x_1,\ldots,x_n}$ from a distribution $P$ with density $p$ supported on $\mathcal{X} = [0,1]^d$. For a given kernel $K_r(x,y) = \frac{1}{r^d}K(\norm{x - y})$, let $G_{n,r}$ denote the random neighborhood graph formed over the samples $X$. Let $L = D - A$ be the corresponding combinatorial Laplacian. In this week's notes, we will show that as long as $r$ is chosen to be sufficiently small, the eigenvalues of $L$ we care about scale at an appropriate rate.

\section{Random Geometric Graph}
In the special case when $K(x) = 1(x \leq r)$ and $G_{n,r}$ is a random geometric graph, we can use results from optimal transport theory to  will prove the following Lemma.
\begin{lemma}
	\label{lem:higher_order_eigenvalue_tail_bound}
	Let $\kappa = n^{2d/(4s + d)}$. Suppose $p$ is uniformly bounded away from $0$ and $\infty$ on its support,
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Then there exists a constant $c$ such that for any $r \leq n^{-4/((2+d)(4s + d))}$, the eigenvalue $\lambda_{\kappa}$ is lower bounded,
	\begin{equation}
	\label{eqn:higher_order_eigenvalue_tail_bound}
	\lambda_{\kappa} \geq c\kappa^{2/d}nr^{2 + d}
	\end{equation}
	with probability $1 - o(n^{-1})$ as $n \to \infty$.
\end{lemma}
Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} establishes that eigenvalues of the graph Laplacian, suitably normalized, scale comparably to the eigenvalues of the continuum Laplacian operator $\mathcal{D}f := \textrm{div}(\nabla f)$. Along with an upper bound on the graph Sobolev semi-norm $R_{s,n}(f)$ (established in the \textit{12.20.19 Notes}), this allows us to argue that if $f$ is a smooth function with large $\mathcal{L}_2$ norm, then the projection of a function $f$ onto the eigenvectors $v_1,\ldots,v_{\kappa}$ is still quite large (measured in empirical norm); or, equivalently, the bias induced by projection onto these eigenvectors is small.

\section{Higher-order kernels.}
When $K$ is the uniform kernel, the aforementioned upper bound on the graph Sobolev semi-norm holds only when $s = 1$ or $s = 2$. For larger values of $s$, we require that $K$ be a higher-order kernel, that is that $K$ satisfy
\begin{equation*}
\int K(x) \,dx = 1,~~ \int x^{\ell}K(x) \,dx = 0~~\textrm{for $\ell = 1,\ldots,s-1$}
\end{equation*}
It seems difficult to extend the techniques used to prove Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} to this setting. We will therefore use a different approach to prove the following Lemma.
\begin{lemma}
	Let $\kappa = n^{2d/(4s + d)}$. Suppose $p$ is uniformly bounded away from $0$ and $\infty$ on its support,
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Let $K$ satisfy the following conditions:

	\textcolor{red}{TODO:} Specify conditions.

	Then there exists a constant $c$ such that the eigenvalue $\lambda_{\kappa}$ is lower bounded,
	\begin{equation}
	\label{eqn:higher_order_eigenvalue_tail_bound2}
	\lambda_{\kappa} \geq c\kappa^{2/d}nr^{2 + d}
	\end{equation}
	with probability $1 - o(n^{-1})$ as $n \to \infty$.
\end{lemma}

@@

The eigenvalues of lattices and complete graphs are known to satisfy, respectively
\begin{equation*}
\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
\end{equation*}
and by standard facts regarding the eigenvalues of tensor product graphs (\textcolor{red}{reference}), we have that the spectrum $\Lambda(H)$ satisfies
\begin{equation*}
\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
\end{equation*}
For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
\begin{align*}
\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
\end{align*}
where the last inequality can be verified by a quick calculation in light of $\kappa = n^{2d/(4s + d)}$ and $r \leq n^{-4/((2+d)(4s + d))}$. Since we've already shown $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ with probability $1 - o(n^{-1})$, this completes the proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}.

Let $D_{\min} = \min_{i \in [n]} D_{ii}$ be the minimum degree of any sample $x \in X$ in the graph $G_{n,r}$. We have that
\begin{equation*}
\lambda_{\kappa}(L) \geq D_{\min} - \lambda_{n - \kappa}(A)
\end{equation*}
We

\end{document}
